### 2023-03-17 code/analysis_code/statistical_analysis.qmd

---
title: "Statistical Analysis"
author: "Sara Benist"
date: "2023-03-17"
output: html_document
---

# Processing script with R script code

This Quarto file runs the statistical analysis code from the `statistical_analysis.r` script. The processed data is used for model fitting.

To start, the needed packages are loaded, and the `processeddata.rda` is loaded from the `processed_data` folder.

# Setup
The `statistical_analysis.r` script needs to be indicated as the file from which to pull code chunks from. 
```{r, cache=FALSE}
knitr::read_chunk('statistical_analysis.R')
```

# Load packages and data
The packages and data are loaded.
```{r, packages, message = FALSE, warning = FALSE}
```

```{r, loaddata}
```
# Model fitting
Now that the data has been explored, I would like to start looking at models for the main outcomes and predictors of the data set. First, we will focus on linear models of the outcomes TB incidence, TB mortality, and TB prevalence. I will be fitting two models using the full data set and only the high burden countries.

## Linear model fitting for TB Incidence

```{r, fitmodel1}
```
The high burden model has the linear equation of `TB incidence = 194 + 126(Male)` where male would equal 1. The full data model has the linear equation of `TB incidence = 82.6 + 48(Male)`. Since many of the countries in the full data set have very low incidence, the coefficients for the intercept and subgroup variable are much lower. The model has more data points with low values.

### Prediction model for TB Incidence
We can also try to predict the approximate TB incidence for a group. For this example, we will include subgroup and setting as predictors.  
```{r, pred1}
```
With the prediction model including setting, there are a lot of variables in the model. The prediction would have a 1 for the country of interest and 0 for all other setting variables. For example, the prediction for a male in Central African Republic would appear as `TB incidence = 126(1)+189(1)`.

As an example, I will predict the incidence rate for males and females in India. The code for prediction modeling is based off of the `tidymodels` guidance. The plot shows the incidence rate and confidence intervals for males (249.3660; CI = 139.68336, 359.0486) and females (123.1038; CI = 13.42112, 232.7864) in India. This can be repeated with the high burden countries.

## Linear model fitting for TB Mortality

```{r, fitmodel2}
```
For high burden countries, the linear equation is `TB mortality = 23.6 + 14.9(Male)`; for the full data model, the linear equation is `TB mortality = 9.98 + 6.43(Male)`.

## Linear model fitting for TB Prevalence

```{r, fitmodel3}
```
For high burden countries, the linear equation is `TB Prevalence = 399 + 151(Urban)` with urban residences receiving a 1 and rural residences receiving a 0. The full data model has an equation of `TB Prevalence = 410 + 95.5(Urban)`. The difference between the high burden and full data models is lower than the other outcomes, most likely due to the long duration of TB or less data points available with the full data model.

## Linear model fitting for BCG Coverage
Now we can look at some of the other indicators in the data set. I will only be looking at the indicators that had variation between subgroups. 
```{r, fitmodel4}
```
The BCG coverage has 12 variables that produce a linear equation of `BCG coverage % = 88.4 + 0.363(Male) -10.6(no education) + 0.169(primary education) + 6.66(higher education) - 6.46(quintile 1) - 1.6(quintile 2) +1.06(quintile 3) + 4.50(quintile 4) + 7.01(quintile 5) - 2.12(rural) + 5.10(Urban)`. The full model has an equation of `BCG coverage % = 88.3 + 0.245(Male) -9.44(no education) + 0.958(primary education) + 6.42(higher education) - 6.40(quintile 1) - 1.68(quintile 2) + 0.773(quintile 3) + 3.84(quintile 4) + 6.41(quintile 5) - 2.08(rural) + 4.75(Urban)`. The two models are overall pretty similar and can be used in the same way as the prediction model (entering in 1 or 0 depending on categorization).

## Linear model fitting for Catastrophic costs

```{r, fitmodel5}
```
The high burden model has an equation of `% catastrophic costs = 87.3 - 40.4(drug-suseptible TB) - 18.6(quintile 1) - 38.8(quintile 2) - 44.6(quintile 3) - 47.4(quintile 4) - 58.5(quintile 5)`. The full model data has an equation of `% catastrophic costs = 83.5 - 35.1(drug-suseptible TB) - 11.1(quintile 1) - 30.9(quintile 2) - 38.4(quintile 3) - 41.4(quintile 4) - 54.9(quintile 5)`.
## Linear model fitting for case detection rate

```{r, fitmodel6}
```
For high burden countries, the equation is `case detection rate % = 37.9 + 19.1(15+ or older) + 16.7(female) + 16.4(male)`. The full data model has an equation of `case detection rate % = 43.8 + 18.9(15+ or older) + 15.3(female) + 16.9(male)`.

## Linear model fitting for prevalence to notification ratio

```{r, fitmodel7}
```
The high burden model has an equation of `prevalence to notification ratio in years = 1.73 + 0.664(male)`, and the full model has an equation of `prevalence to notification ratio in years = 1.73 + 0.701(male)`.

## Linear model fitting for attitudes

```{r, fitmodel8}
```
The attitude question regarding TB was about the desire to keep a diagnosis hidden. Since males are more likely to be diagnosed with TB, I focused on the male attitude and knowledge. For high burden countries, the equation is `% wanting to keep TB secret = 27.6 - 5.27(20-34 years) - 8.21(35-49 years) - 3.53(no education) - 2.71(primary education) - 7.23(higher education) - 5.01(quintile 1) - 5.23(quintile 2) - 4.97(quintile 3) - 4.38(quintile 4) - 5.86(quintile 5) - 4.91(rural) - 6.27(urban)`. For the full model, the equation is `% wanting to keep TB secret = 24.7 - 4.94(20-34 years) - 8.14(35-49 years) - 4.32(no education) - 3.08(primary education) - 6.41(higher education) - 4.44(quintile 1) - 5.24(quintile 2) - 5.18(quintile 3) - 4.59(quintile 4) - 6.20(quintile 5) - 4.66(rural) - 5.99(urban)`.

## Linear model fitting for knowledge

```{r, fitmodel9}
```
The knowledge question asked about if TB is spread through cough. We will again focus on male knowledge.

The high burden countries had an equation of `% responding TB is spread through cough = 65.9 + 4.83(20-34 years) + 6.19(35-49 years) - 9.91(no education) - 2.81(primary education) + 12.6(higher education) - 6.01(quintile 1) - 0.464(quintile 2) + 2.32(quintile 3) + 7.50(quintile 4) + 14.6(quintile 5) + 0.419(rural) + 11.8(urban)`. The full model had an equation of `% responding TB is spread through cough = 64.6 + 6.17(20-34 years) + 7.29(35-49 years) - 8.55(no education) - 1.66(primary education) + 12.8(higher education) - 3.88(quintile 1) + 0.028(quintile 2) + 3.03(quintile 3) + 8.18(quintile 4) + 15.8(quintile 5) + 1.28(rural) + 12.8(urban)`.

I would like to try and create a dataset using the high burden countries in order to predict the main outcomes of TB incidence and mortality from the indicators (BCG Coverage, knowledge, attitudes, etc). Right now, the data is not set up to complete this analysis, but I will try to find a way to do this.